A recipe for extremely reproducible enrichment analysis (and why we need it)

### 2023-12-04 manuscript/manuscript.Rmd
ack
letter
mandates

(*) Corresponding author: m.ziemann@deakin.edu.au

@@

Unreliable and irreproducible research is a significant problem that wastes resources and risks
undermining the public perception of science.
Previous work has highlighted that published enrichment analyses frequently suffer from statistical
and reporting flaws.
We sought to determine whether this translates into irreproducibility by examining whether the
findings of 20 open-access articles published in 2019 describing enrichment analysis with the
popular DAVID suite were reproducible.
We find that only three articles exhibited a high degree of concordance, while seven exhibited major
discrepancies, which we mainly ascribe to deficiencies in methodological reporting.
As the tool version used is no longer available, all articles using this tool pre-2021 (~20,800 studies)
including this sample of 20 articles are no longer able to be reproduced with the original tools.
Based on this, we suggest that results from web-based tools without long-term preservation features should
not be included in scientific publications due to the threat of link decay and short reproducibility horizon.
We advocate for the adoption of extremely reproducible research workflows, and we provide a detailed protocol
for how to achieve it using a combination of best practices including literate programming, version control,
containerisation, documentation and persistent sharing of data and software.

@@

We thank Ms Megan Soria (Deakin University) for discussions on the concept and manuscript.
This research was supported by use of the Nectar Research Cloud, a collaborative Australian research
platform supported by the NCRIS-funded Australian Research Data Commons (ARDC).

>>>>

(*) Corresponding author: mark.ziemann@burnet.edu.au

@@

Unreliable and irreproducible research is a significant problem that wastes resources and risks
undermining the public perception of science.
Previous work has highlighted that published enrichment analyses frequently suffer from statistical
and reporting flaws.
We sought to determine whether this translates into irreproducibility by examining whether the
findings of 20 open-access articles published in 2019 describing enrichment analysis with the
popular DAVID suite were reproducible.
We find that only three articles exhibited a high degree of concordance, while seven exhibited major
discrepancies, which we mainly ascribe to deficiencies in methodological reporting.
As the tool version used is no longer available, all articles using this tool pre-2021 (~20,800 studies)
including this sample of 20 articles are no longer able to be reproduced with the original tools.
Based on this, we suggest that results from web-based tools without long-term preservation features should
not be included in scientific publications due to the threat of link decay and short reproducibility horizon.
Relying exclusively on webtools for analysis may also be in breach of institutional and funder data management mandates.
We advocate for the adoption of extremely reproducible research workflows, and we provide a detailed protocol
for how to achieve it using a combination of best practices including literate programming, version control,
containerisation, documentation and persistent sharing of data and software.

@@

We thank Ms Megan Soria (Deakin University) for discussions on the concept and manuscript.
This research was supported by use of the Nectar Research Cloud, a collaborative Australian research
platform supported by the NCRIS-funded Australian Research Data Commons (ARDC).
The authors gratefully acknowledge the contribution to this work of the
Victorian Operational Infrastructure Support Program received by the Burnet Institute.
